Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2020-01-16 writeups/Su_disk_disp.tex
stash, finished all S2

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\renewcommand*{\sectionautorefname}{Section}
% \renewcommand*{\subsectionautorefname}{Subsection}
\maketitle

@@

The paper is organized as follows. In \autoref{s:eq}, we review relevant
spin-orbit dynamics and present the equations studied in the rest of the paper.
In \autoref{s:ad} and \autoref{s:nonad}, we study the evolution of the system
when the disk either dissipates adiabatically or non-adiabatically respectively.
Exhaustive yet simple analytical arguments are able to explain completely the
numerical results in both limits. We discuss potential consequences in
\autoref{s:disc}. Our primary physical results consist of
\autoref{fig:ad_ensemble} in the adiabatic limit and
\autoref{fig:3_ensemble_05_15} in the non-adiabatic limit. Longer mathematical
calculations are relegated to the appendicies, including a leading-order
estimate of final planetary obliquities given small initial spin-disk
misalignment angles in \autoref{s:ad_approx}.

@@

Much of this treatment parallels that of \citet{anderson2018teeter} and
\citet{millholland_disk}. We consider an oblate star of mass $M_\star$ hosting a
planet of mass $M_{\rm p}$, radius $R_{\rm p}$, at semimajor axis $a_{\rm p}$,
and with spin angular frequency $\Omega_{\rm p}$, and a protoplanetary disk of
mass $M_{\rm d}$ at some characteristic distance $r_{\rm d}$ from the host star
\citep[see][for a power-law disk profile]{millholland_disk}. Denote $\hat{s}
\equiv \bm{S} / S$ spin unit vector of planet, $\hat{l} \equiv \bm{L} / L$
orbital angular momentum unit vector of planet, and $\hat{l}_{\rm d} \equiv
\bm{L}_{\rm d} / L_{\rm d}$ angular momentum of the surrounding disc. We
approximate $S \ll L \ll L_{\rm d}$, so $\hat{l}_{\rm d}$ is approximately
constant and $\hat{l}$ experiences negligible backreaction torque from
$\hat{s}$. The quadrupole-order secular equations of motion for $\hat{s}$,
$\hat{l}$ and $\hat{l}_{\rm d}$ then become very simple
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{\rm sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}},\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{\rm ld}\p{\hat{l} \cdot \hat{l}_{\rm d}}
        \p{\hat{l} \times \hat{l}_{\rm d}}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm
        p}} \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &\equiv \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
        p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In \autoref{eq:wsl}, $k_{\rm qp}$ and $k_{\rm }$ are related to the planet's
quadrupole moment and moment of inertia respectively \citep[see][]{lai2018}, and
$n \equiv \sqrt{GM_\star/a_{\rm p}^3}$ is the planet's orbital mean motion. The
three angles formed by the three vectors are denoted
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{\rm sd} &\equiv \arccos \hat{s} \cdot \hat{l}_{\rm d},\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_{\rm d}.
\end{align}

Let us transform to frame rotating about $\hat{l}_{\rm p}$ with frequency $g
\equiv -\omega_{\rm ld}\cos I < 0$, then $\hat{l}$ is fixed. In this frame,
$\hat{l}_{\rm p}$ also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_{\rm d}},
\end{equation}
where $\alpha \equiv \omega_{\rm sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta \equiv{}& \frac{\abs{g}}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}}{0.001 M_\odot}}
            \p{\frac{R_{\rm p}}{R_J}}^{-3}
        &\p{\frac{M_{\rm d}}{0.01 M_{\odot}}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^{9/2}
            \p{\frac{r_{\rm d}}{30 \;\mathrm{AU}}}^{-3}
            \p{\frac{M_\star}{1M_{\odot}}}^{-3/2}
            \cos I,\\
    \tau \equiv{}& \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_{\rm d}}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_{\rm d}(t) = M_{\rm d}(0)e^{-t/t_{\rm d}}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_{\rm d}}$ as well. We nondimensionalize:
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_{\rm d}},\\
        ={}& 0.015 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}/M_\star}{0.001}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^3\nonumber\\
        &\p{\frac{R_{\rm p}}{R_J}}^{-3}
            \p{\frac{\Omega_{\rm p}}{2\pi/\p{10\;\mathrm{days}}}}
            \p{\frac{t_{\rm d}}{10\;\mathrm{Myr}}}^{-1},\\
    \rd{\eta}{\tau} ={}& -\epsilon \eta.\label{eq:deta_dt}
\end{align}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

In the next two subsections, we further discuss theoretical background relevant
to our analysis of the evolution of this system. They are not essential to
reproducing the results of \autoref{s:ad} and \autoref{s:nonad}.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(see \autoref{s:local_dynamics}). When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$. Note that
    $\theta \in [-\pi, \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}. The vertical dashed
    line indicates $\eta_{\rm c}$, where CS1 and CS4 merge and annihilate, and the
    horizontal dashed lines indicate $\theta = I, 180^\circ - I$ the asymptotic
    behavior for CS2, 3 for $\eta \gg \eta_{\rm c}$.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. In
\autoref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs via a contour plot of the Hamiltonian of the system
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}}.\label{eq:H}
\end{equation}
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$/$III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    \autoref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_{\rm c}$.
    The three zones, divided by the separatrix, are labeled. The Cassini states
    are labeled and have the same colors as \autoref{fig:cs_locs}. The interior
    of the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the points in phase space that flow into
    zone $II$ when evolved forward in time; this is the shaded region in the top
    left plot, bounded by the dotted black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},\\
    \end{align}
\end{subequations}
the areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$, but a natural
extension exists: evolve an initial point $p$ under adiabatic decrease of $\eta$
until the separatrix appears at $\eta = \eta_{\rm c}$, then identify $p$ with
the zone it is in at $\eta_{\rm c}$. Since phase space area is conserved under
adiabatic evolution, this implies $A_i\p{\eta > \eta_{\rm c}} = A_i(\eta_{\rm
c})$. The boundary between these extended zones is denoted by the dashed black
line in the top left panel of \autoref{fig:eq_1contours}, where otherwise no
separatrix should exist since $\eta > \eta_{\rm c}$.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by \autoref{se:area_ward}. Dotted lines correspond to small
    $\eta$ approximations discussed in \autoref{s:ad_approx}. Dashed lines for
    $\eta > \eta_{\rm c}$ effective values of $A_{\rm II}, A_{\rm III}$ for
    $\eta > \eta_{\rm c}$, denoting the area of points that would flow into
    either area under adiabatic decrease of $\eta$ from $\eta > \eta_{\rm
    c}$.}\label{fig:eq_areas}
\end{figure}

@@

As is shown in \autoref{ss:eigens}, the libration frequency about CS2 is slower
than that about CS1 or CS3. As such, it is the slowest characteristic frequency
in the system. The weak adiabaticity criterion can be recast as requiring that
$\eta$ vary more slowly than this libration period; this is exactly the
criterion used in \citep{ward2004II,millholland_disk}. This constraint can be
computed
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}} = \epsilon &\lesssim
            \frac{1}{T_{\rm lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta\sin I \sin \theta_2
            \p{1 + \eta \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{align}
where $\theta_2$ is the obliquity at CS2. This formula differs somewhat from
that presented in \citet{millholland_disk}, where the $\csc^3\theta_2$ term is
neglected and a square root is missing.

For $I = 5^\circ$, $\eta \sim \eta_c$ we compute $\epsilon \lesssim 0.085$ is
the requirement for weak adiabaticity. Since our criterion is only a weak
constraint on adiabaticity, we use $\epsilon = 3 \times 10^{-4}$ in our
``adiabatic'' simulations. We explore the consequences of increasingly
nonadiabatic evolution in \autoref{ss:transition}.

@@

We evenly sample $101$ values of $\theta_{\rm sd, i}$, and for each $\theta_{\rm
sd, i}$ value, we pick $101$ $\hat{s}$ initial conditions approximately on the
ring fixed by $\theta_{\rm sd, i}$\footnote{For each $\theta_{\rm sd, i}$ we
choose an initial condition on the libration cycle going through $\p{\theta_2 +
\theta_{\rm sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates
of CS2), such that all initial conditions for a given $\theta_{\rm sd, i}$
enclose the same initial phase space; see \autoref{sss:a_evo}.}. To be concrete,
we choose $\eta_i = 10\eta_{\rm c}$. For each initial condition, we evolve
Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt} until $\eta = 10^{-5}$. For
such small $\eta$, $\hat{s}$ precesses about $\hat{l}$ with uniform final
obliquity $\theta_{\rm f}$. This transfer function $\theta_{\rm f}\p{\theta_{\rm
sd, i}}$ is our primary result, and is shown for $I = 5^\circ, 10^\circ,
20^\circ$ in \autoref{fig:ad_ensemble}, \autoref{fig:3_ensemble_10_35}, and
\autoref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the simulation, which are in excellent agreement with the colored tracks.
These colored tracks are calculated semi-analytically using methods explained in
the following section.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{\rm f}\p{\theta_{\rm sd, i}}$, overlaid with
    semi-analytic predictions of the $\theta_{\rm  f}$ for each of the four
    nontrivial dynamical tracks in colored lines. Black dotted lines correspond
    to small-$\theta_{\rm sd, i}$ predictions derived in \autoref{s:ad_approx}.
    Bottom: Semi-analytic probabilities of each of the dynamical tracks for each
    $\theta_{\rm sd, i}$. The five regimes of $\theta_{\rm sd, i}$ values
    demarcated by the vertical dashed black lines correspond to the five regimes
    of $A_i$ calculated in \autoref{ss:zone_transitions}. Black dotted lines
    correspond to simple analytic estimates given in
    \autoref{s:ad_approx}.}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:3_ensemble_10_35} but for $I = 20^\circ$. The
    $III \to II \to I$ track is very short since the maximum $A_{\rm II}$ is very
    near $\eta_{\rm c}$ for $I = 20^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

>>>>

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

@@

The paper is organized as follows. In Section~\ref{s:eq}, we review relevant
spin-orbit dynamics and present the equations studied in the rest of the paper.
In Sections~\ref{s:ad} and~\ref{s:nonad}, we study the evolution of the system
when the disk either dissipates adiabatically or non-adiabatically respectively.
Exhaustive yet simple analytical arguments are able to explain completely the
numerical results in both limits. We discuss potential consequences in
Section~\ref{s:disc}. Our primary physical results consist of
Fig.~\ref{fig:ad_ensemble} in the adiabatic limit and
Fig.~\ref{fig:3_ensemble_05_15} in the non-adiabatic limit. Longer mathematical
calculations are relegated to the appendicies, including a leading-order
estimate of final planetary obliquities given small initial spin-disk
misalignment angles in Appendix~\ref{s:ad_approx}.

@@

Much of this treatment parallels that of \citet{anderson2018teeter} and
\citet{millholland_disk}. We consider a star of mass $M_\star$ hosting an oblate
planet of mass $M_{\rm p}$, radius $R_{\rm p}$, at semimajor axis $a_{\rm p}$,
and with spin angular frequency $\Omega_{\rm p}$, and a protoplanetary disk of
mass $M_{\rm d}$ at some characteristic distance $r_{\rm d}$ from the host star
\citep[see][for a power-law disk profile]{millholland_disk}. Denote $\hat{s}
\equiv \bm{S} / S$ spin unit vector of planet, $\hat{l} \equiv \bm{L} / L$
orbital angular momentum unit vector of planet, and $\hat{l}_{\rm d} \equiv
\bm{L}_{\rm d} / L_{\rm d}$ angular momentum of the surrounding disc. We
assume $S \ll L \ll L_{\rm d}$, so $\vec{l}_{\rm d}$ is approximately
constant and $\vec{l}$ experiences negligible backreaction torque from
$\hat{s}$. The equations of motion for $\hat{s}$, $\hat{l}$ and $\hat{l}_{\rm
d}$ are
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{\rm sl} \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        \equiv \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}},\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{\rm ld}\p{\hat{l} \cdot \hat{l}_{\rm d}}
            \p{\hat{l} \times \hat{l}_{\rm d}}
        \equiv -g\p{\hat{l} \times \hat{l}_{\rm d}},
            \label{eq:dldt}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &= \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm
        p}} \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &= \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
        p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In Eq.~\eqref{eq:wsl}, $k_{\rm qp}$ and $k_{\rm p}$ are related to the planet's
rotation-induced quadrupole moment and moment of inertia respectively
\citep[see][]{lai2018}, and $n \equiv \sqrt{GM_\star/a_{\rm p}^3}$ is the
planet's orbital mean motion. The three angles formed by the three vectors are
denoted
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{\rm sd} &\equiv \arccos \hat{s} \cdot \hat{l}_{\rm d},\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_{\rm d}.
\end{align}
Following standard notation \citep[e.g.][]{colombo1966,peale1969,ward_jupiter},
we have defined $\alpha \equiv \omega_{\rm sl}$ and $g = -\omega_{\rm ld}\cos
I$.

We can combine Eqs.~\eqref{eq:dsdt}--\eqref{eq:dldt} into a single equation by
transforming into a frame rotating about $\hat{l}_{\rm d}$ with frequency $g$.
In this frame, $\hat{l}_{\rm d}$ and $\hat{l}$ are fixed, and $\hat{s}$ evolves
as:
\begin{equation}
    \at{\rd{\hat{s}}{t}}_{\rm rot} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_{\rm d}}.\label{eq:dsdt_rot}
\end{equation}

We define the frequency ratio $\eta$ and dimensionless time $\tau$ as
\begin{align}
    \eta \equiv{}& -\frac{g}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}}{0.001 M_\odot}}
            \p{\frac{R_{\rm p}}{R_J}}^{-3}
        &\p{\frac{M_{\rm d}}{0.01 M_{\odot}}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^{9/2}
            \p{\frac{r_{\rm d}}{30 \;\mathrm{AU}}}^{-3}
            \p{\frac{M_\star}{1M_{\odot}}}^{-3/2}
            \cos I,\\
    \tau \equiv{}& \alpha t.
\end{align}
Then Eq.~\eqref{eq:dsdt_rot} becomes
\begin{equation}
    \at{\rd{\hat{s}}{t}}_{\rm rot} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_{\rm d}}. \label{eq:dsdt_base}
\end{equation}

Throughout this paper, we consider $\alpha$ constant, but allow $g$ to vary in
time. In the dispersing disk scenario of \citet{millholland_disk}, $\abs{g}$
decreases in time due to the decreasing disk mass. We consider a simple
exponential decay model
\begin{equation}
    M_{\rm d}(t) = M_{\rm d}(0)e^{-t/t_{\rm d}},
\end{equation}
with $t_d$ constant.

Tihs implies $\rd{\eta}{t} = -\frac{\eta}{t_{\rm d}}$. In dimenisonless units,
we have
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta\label{eq:deta_dt},
\end{equation}
where
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_{\rm d}},\\
        ={}& 0.015 \p{\frac{k_{\rm p}}{6k_{\rm qp}}}
            \p{\frac{m_{\rm p}/M_\star}{0.001}}
            \p{\frac{a_{\rm p}}{1\;\mathrm{AU}}}^3\nonumber\\
        &\p{\frac{R_{\rm p}}{R_J}}^{-3}
            \p{\frac{\Omega_{\rm p}}{2\pi/\p{10\;\mathrm{days}}}}
            \p{\frac{t_{\rm d}}{10\;\mathrm{Myr}}}^{-1}.
\end{align}
Eqs.~\eqref{eq:dsdt_base} and~\eqref{eq:deta_dt} together constitute our system
of study.

In the next two subsections, we discuss and review the theoretical background
relevant to our analysis of the evolution of the system.

@@

Spin states satisfying $\at{\rd{\hat{s}}{t}}_{\rm rot} = 0$ are referred to as
\emph{Cassini States} (CSs). They require that $\hat{s}$, $\hat{l}$, and
$\hat{l}_{\rm d}$ be coplanar. There are either two or four solutions, depending
on the values of $\eta, I$. They are specified by the obliquity $\theta$ and the
precessional phase of $\hat{s}$ around $\hat{l}$, denoted $\phi$. Following
standard convention and nomenclature (see Fig.~\ref{fig:cs_locs}), CSs 1, 3, 4
have $\phi = 0$ and $\theta < 0$, corresponding to $\hat{s}$ and $\hat{l}_{\rm
d}$ being on opposite sides of $\hat{l}$, while CS2 has $\phi = \pi$ and $\theta
> 0$, corresponding to $\hat{s}$ and $\hat{l}_{\rm d}$ being on the same side of
$\hat{l}$. When $\eta < \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_{\rm c}$, only CSs 2, 3 exist. The CS
obliquities as a function of $\eta$ are shown in Fig.~\ref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ at $I = 5^\circ$.
    Note that $\theta \in [-\pi, \pi]$ is the traditional definition of the
    polar angle. The vertical dashed line indicates $\eta_{\rm c}$, where CS1
    and CS4 merge and annihilate, and the horizontal dashed lines indicate
    $\theta = I$ and $180^\circ - I$, the asymptotic behavior for CSs 2 and 3
    for $\eta \gg \eta_{\rm c}$.}\label{fig:cs_locs}
\end{figure}

Of the four CSs, 1, 2, 3 are stable while 4 is unstable.
Appendix~\ref{s:local_dynamics} gives the libration frequencies and growth rates
respectively near these CSs.

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two
trajectories originating and ending at CS4 are the only two infinite-period
orbits in the phase space. Together, these two critical trajectories are
referred to as the \emph{separatrix} and divide phase space into three zones. In
Eq.~\eqref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs via a contour plot of the Hamiltonian of the system
\begin{align}
    \mathcal{H} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}},\label{eq:H}\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$ and $III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    Eq.~\eqref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_{\rm c}$.
    The three zones, divided by the separatrix, are labeled. The Cassini states
    are labeled and have the same colors as in Fig.~\ref{fig:cs_locs}. The
    interior of the separatrix, shaded in grey, is formally only defined for
    $\eta < \eta_{\rm c}$, but we may identify the points in phase space that
    flow into zone $II$ when evolved forward in time (decreasing $\eta$
    adiabatically); this is the shaded region in the top left panel, bounded by
    the black dotted line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas $\p{\abs{\int \cos \theta \;\mathrm{d}\phi}}$ of the three
zones can be computed analytically \citep{henrard1987,ward2004I}. Defining
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},\\
    \end{align}
\end{subequations}
the areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this implies
$A_i\p{\eta > \eta_{\rm c}} = A_i(\eta_{\rm c})$. The boundary between these
extended zones is denoted by the dashed black line in the top left panel of
Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by Eqs.~\eqref{se:area_ward}. Dotted lines correspond to
    small $\eta$ approximations discussed in Appendix~\ref{s:ad_approx}. Dashed
    lines for $\eta > \eta_{\rm c}$ effective values of $A_{\rm II}, A_{\rm
    III}$ for $\eta > \eta_{\rm c}$, denoting the area of points that would flow
    into either area under adiabatic decrease of $\eta$ from $\eta > \eta_{\rm
    c}$.}\label{fig:eq_areas}
\end{figure}

@@

As is shown in Appendix~\ref{ss:eigens}, the libration frequency about CS2 is slower
than that about CS1 or CS3. As such, it is the slowest characteristic frequency
in the system. The weak adiabaticity criterion can be recast as requiring that
$\eta$ vary more slowly than this libration period; this is exactly the
criterion used in \citep{ward2004II,millholland_disk}. This constraint can be
computed
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}} = \epsilon &\lesssim
            \frac{1}{T_{\rm lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta\sin I \sin \theta_2
            \p{1 + \eta \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{align}
where $\theta_2$ is the obliquity at CS2. This formula differs somewhat from
that presented in \citet{millholland_disk}, where the $\csc^3\theta_2$ term is
neglected and a square root is missing.

For $I = 5^\circ$, $\eta \sim \eta_c$ we compute $\epsilon \lesssim 0.085$ is
the requirement for weak adiabaticity. Since our criterion is only a weak
constraint on adiabaticity, we use $\epsilon = 3 \times 10^{-4}$ in our
``adiabatic'' simulations. We explore the consequences of increasingly
nonadiabatic evolution in Section~\ref{ss:transition}.

@@

We evenly sample $101$ values of $\theta_{\rm sd, i}$, and for each $\theta_{\rm
sd, i}$ value, we pick $101$ $\hat{s}$ initial conditions approximately on the
ring fixed by $\theta_{\rm sd, i}$\footnote{For each $\theta_{\rm sd, i}$ we
choose an initial condition on the libration cycle going through $\p{\theta_2 +
\theta_{\rm sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates
of CS2), such that all initial conditions for a given $\theta_{\rm sd, i}$
enclose the same initial phase space; see Section~\ref{sss:a_evo}.}. To be concrete,
we choose $\eta_i = 10\eta_{\rm c}$. For each initial condition, we evolve
Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt} until $\eta = 10^{-5}$. For
such small $\eta$, $\hat{s}$ precesses about $\hat{l}$ with uniform final
obliquity $\theta_{\rm f}$. This transfer function $\theta_{\rm f}\p{\theta_{\rm
sd, i}}$ is our primary result, and is shown for $I = 5^\circ, 10^\circ,
20^\circ$ in Figs.~\ref{fig:ad_ensemble},~\ref{fig:3_ensemble_10_35},
and~\ref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the simulation, which are in excellent agreement with the colored tracks.
These colored tracks are calculated semi-analytically using methods explained in
the following section.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{\rm f}\p{\theta_{\rm sd, i}}$, overlaid with
    semi-analytic predictions of the $\theta_{\rm  f}$ for each of the four
    nontrivial dynamical tracks in colored lines. Black dotted lines correspond
    to small-$\theta_{\rm sd, i}$ predictions derived in
    Appendix~\ref{s:ad_approx}. Bottom: Semi-analytic probabilities of each of
    the dynamical tracks for each $\theta_{\rm sd, i}$. The five regimes of
    $\theta_{\rm sd, i}$ values demarcated by the vertical dashed black lines
    correspond to the five regimes of $A_i$ calculated in
    Section~\ref{ss:zone_transitions}.}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as Fig.~\ref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as Fig.~\ref{fig:3_ensemble_10_35} but for $I = 20^\circ$. The
    $III \to II \to I$ track is very short since the maximum $A_{\rm II}$ is very
    near $\eta_{\rm c}$ for $I = 20^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}
